---
title: "Fractal pattern simulation"
author: "Nicolas Sutton-Charani"
date: "2023-11-20"
output: 
  html_document:
    code_folding: hide
    fig_width: 6 
    fig_height: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nonlinearTseries); library(ggplot2); library(ggpubr); library(rmarkdown)
```

## Context

The DFA is suposed to compute the level of fractality of a signal but it is done in terms of fluctuation, i.e. in terms of deviation to local (linear) trends.

## Objective

Extending the DFA to signal **pattern** fractality, i.e. the ability of retrieving distributionnal patterns at different terms (or scale).

## Method

Simulate pattern-fractality signals of different levels of fractality and observe the DFA behavior. To do so, the following steps:

* simulate a gaussian signal
* compute its histogram $\rightarrow$ the initial pattern
* fractal convolution of the pattern several times:
  + linear interpolate each pattern segment to get segment granularity = pattern granularity
  + add a deweighted version of the pattern to each segment pattern: 
  
  new segment $\leftarrow$ segment + pattern $\times \frac{1}{\text{fractal factor}}$

```{r pattern, fig.align='center'}
set.seed(123)#, out.width="50%", out.height="10%"
sig <- rnorm(100)
nbreaks <- 5
h <- hist(sig, breaks=nbreaks, plot=F)
pattern <- h$density

p_initial_sig <- ggplot(data.frame(x=1:length(sig), 
                                   y=sig), aes(x,y)) + 
  geom_line() +
  xlab("") + ylab("") +
  ggtitle("initial signal") +
  theme_bw() +
  theme(axis.ticks.x=element_blank(), axis.ticks.y=element_blank(), 
        axis.text.x=element_blank())
p_initial <- ggplot(data.frame(x=h$mids, 
                               y=pattern), aes(x,y)) + 
  geom_line() +
  xlab("") + ylab("") +
  ggtitle("initial pattern") +
  theme_bw() +
  theme(axis.ticks.x = element_blank(), axis.ticks.y = element_blank(),
        axis.text.y=element_blank())
ggarrange(p_initial_sig, p_initial)
```

```{r several_convolutions, fig.align='center'}

fractal_factor <- 5
n <- length(pattern)

final_sig <- pattern
alphas <- c()
sigs <- list()
sigs[[1]] <- pattern
plots <- list()
plots[[1]] <- p_initial
i_window=4
n_convol <- 10
for (i_convol in 1:n_convol){
  new_sig <- final_sig * 1/fractal_factor + 
    approx(x=1:length(final_sig), 
           y=final_sig, 
           n=fractal_factor*length(final_sig))$y
  new_sig <- new_sig * (max(pattern)/max(new_sig))
  final_sig <- new_sig
  alpha <- as.numeric(estimate(dfa(final_sig, do.plot=F)))
  cat(alpha, " ")
  plots[[i_convol+1]] <- ggplot(data.frame(x=1:length(final_sig), y=final_sig), aes(x,y)) +
    geom_line() +
    ggtitle(paste(i_convol, "convolutions -> alpha =", round(alpha, 3)))+
    theme_bw()
  alphas <- c(alphas, alpha)
}

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], 
          plots[[5]], plots[[6]], plots[[7]], plots[[8]], 
          plots[[9]], plots[[10]], plots[[11]], NULL,
          nrow=3, ncol=4)
```

